\LARGE How coexistence mechanisms mediate temporal stability

### 2014-10-16 modelNotes.Rnw
updated model notes
udpated doc

\documentclass[12pt]{article}
\usepackage{graphics}
\usepackage{graphicx, verbatim}
\usepackage{amsmath}
% \usepackage[T1]{fontenc}
% \usepackage[utf8]{inputenc}
\usepackage{authblk}
\textwidth=6.2in
\textheight=8.5in
%\parskip=.3cm
\oddsidemargin=.1in
\evensidemargin=.1in
\headheight=-.3in

@@

\subsection{Temporal storage effect}

>>>>

\documentclass[12pt]{article}
\usepackage{graphics}
\usepackage{graphicx, verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
% \usepackage[T1]{fontenc}
% \usepackage[utf8]{inputenc}
\usepackage{authblk}
\textwidth=6.2in
\textheight=8.5in
%\parskip=.3cm
\oddsidemargin=.1in
\evensidemargin=.1in
\headheight=-.3in

@@

\subsection{Temporal storage effect}
We start with the continuous-time model

\begin{align}
\frac{\text{d}D_{i}}{\text{d}t} &= c_{i}b_{i}N_{i} - g_{i}D_{i} - m_{D,i}D_{i}\\
\frac{\text{d}N_{i}}{\text{d}t} &= f(R,N_{i}) + g_{i}D_{i} - c_{i}b_{i}N_{i} - m_{N,i}N_{i}\\
\frac{\text{d}R_{i}}{\text{d}t} &= a(S - R) - \sum\limits_{i=1,2}f(R,N_{i})
\end{align}
where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $B$ is the living biomass (fast-growing, shorter-lived) state, $b$ is the "birth" rate of new biomass in the dormant state, $c$ is some factor related to the cost of producing the dormant state, $g$ is the time-fluctuating activation rate of dormant biomass to the live biomass state, and $m$s are biomass loss rates. The growth rate of living biomass is a resource-dependent function, $f(R,N) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$. For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent.

To make this a "storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years. Our model meets condition 1 because we include a dormant stage with very low death rates. We satisfy condition 2 with model whenever $g$ is not perfectly correlated between species. Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years. So long as resource uptake rates are equal ... oops we have a problem!
